\newglossaryentry{strength}
{
    name=strength,
    description={Attribute describing the physical strength of a character.}
}

\newglossaryentry{agility}
{
    name=agility,
    description={Attribute describing the nimbleness and flexibility of a
        character.}
}

\newglossaryentry{critical-confirmation}
{
    name={critical confirmation},
    description={After scoring a critical threat, the attacker can make a
    critical confirmation roll before rolling for damage. If the natural roll
    falls within the attacker's critical confirmation range, the critical
    threat is upgraded to a critical hit.}
}

\newglossaryentry{critical-confirmation-range}
{
    name={critical confirmation range},
    description={Every character has a critical confirmation range. The
    critical confirmation range is 20 by default, but can be increased through
    various means. The critical confirmation range is the range of natural
    rolls on a critical confirmation that upgrade a critical threat to a
    critical hit.}
}

\newglossaryentry{critical-damage-multiplier}
{
    name={critical damage multiplier},
    description={Every weapon has a critical damage muliplier. The critical
    damage multiplier is the amount damage is multiplied by when a critical
    hit is scored. Damage is multiplied after damage is rolled and the damage
    formula is calculated.}
}

\newglossaryentry{critical-hit}
{
    name={critical hit},
    description={A critical hit is the result of a successful critical
    confirmation roll, and replaces the effects of a critical threat. An attack
    which is a critical hit is an automatic hit, bypasses armor, and deals
    damage multiplied by the weapon's critical damage multiplier. Damage is
    multiplied after it is rolled and the damage formula is calculated.}
}

\newglossaryentry{critical-threat}
{
    name={critical threat},
    description={A critical threat is scored when the natural roll of an attack
    with a weapon falls within that weapon's critical threat range. A critical
    threat is an automatic hit (but does not necessarily bypass armor), and
    allows the attacker to make a critical confirmation roll.}
}

\newglossaryentry{critical-threat-range}
{
    name={critical threat range},
    description={Every weapon has a critical threat range. The critical threat
    range is the range of natural rolls on an attack that scores a critical
    threat.}
}

\newglossaryentry{dark-vision}
{
    name={dark vision},
    description={A creature with this ability has all the effects of night
        vision. In addition, it can see in darkness within 60 ft. of its
        position in monochrome as if it was dim light.}
}

\newglossaryentry{natural-roll}
{
    name={natural roll},
    description={A natural roll is the number displayed on the die/dice after
    any check or roll is made. Alternatively, it is the result of a die/dice
    roll with no modifiers or penalties applied whatsoever (this definition
    is more useful when using a digital or virtual dice roller).}
}

\newglossaryentry{natural-x}
{
    name={natural X},
    description={A natural roll of a particular number is often described as a
    natural X, where X is the number that was rolled. For example, a roll of
    1d20 where the die displays a 20 could be described as a "natural 20".}
}

\newglossaryentry{night-vision}
{
    name={night vision},
    description={A creature with this ability experiences dim light as bright
        light. In addition, it experiences darkness at the periphary of light
        sources with an origin as dim light up to a distance from the origin 
        twice to the actual dim light radius.}
}

\newglossaryentry{slightly-small}
{
    name={slightly small},
    description={A creature with this ability does not automatically gain the
        small size category. As long as it is medium or larger, it gains +1
        DEX, -1 STR, -1 END and -5 speed.}
}

\newglossaryentry{sunlight-sensitivity}
{
    name={sunlight sensitivity},
    description={A creature with sunlight sunsitivity gains -2 DEX and -2 SEN
        in sunlight. Additionally, it becomes blinded for 1 round when abruptly
        entering sunlight.}
}

\newglossaryentry{slight-sunlight-sensitivity}
{
    name={slight sunlight sensitivity},
    description={A creature with slight sunlight sensitivity gains -1 DEX and
        -1 SEN in sunlight.}
}

\newglossaryentry{small-tusks}
{
    name={Small Tusks},
    description={A creature with small tusks has small tusks as a natural
        weapon instead of teeth. These deal 1d4 + STR piercing damage.}
}
