\chapter{Details}

\begin{multicols*}{2}
    This chapter covers any remaining details you need to calculate to finish
    making a character.

    \section{Encumbrance}\index{encumbrance}
    A character's encumbrance threshold\index{encumbrance threshold} is 10 + STR.

    For each multiple of the encumbrance threshold a character exceeds, that
    character has 1 encumbrance. For example, if a character has an encumbrance
    threshold of 10 and is carrying exactly 30 lb. of equipment, they have an
    encumbrance of 2. If they pick up a 1 lb. item, they are carrying 31 lb. and
    have an encumbrance of 3.

    A character's encumbrance is used as a penalty on certain mobility-related
    skills, chief of which is dodge. For a full list of penalised skills, see the
    skills chapter (pg. \ref{skills}).

    A character has their base move speed reduced by 5 ft. for every 3
    encumbrance. Characters with exotic base move speeds, however, have their
    base move speed reduced according to table \ref{tab:encumbered-move}.
    A character's base move speed can only be reduced as far as 5 ft. in this
    way.

    A character whose encumbrance encumbrance is 10 or more cannot walk
    and cannot use skills which are penalised by encumbrance.

    \begin{center}
        \unclassedrowcolors
        \begin{tabularx}{0.5\textwidth}{X l}
            \textbf{Base move speed} & \textbf{Reduction} \\
            0 - 45 ft. & -5 ft. \\
            50 - 75 ft. & -10 ft. \\
            80 - 105 ft. & -15 ft. \\
            110 - 135 ft. & -20 ft. \\
            etc. & etc. \\
        \end{tabularx}
        \captionof{table}{Encumbered Move Speeds}
        \label{tab:encumbered-move}
    \end{center}
\end{multicols*}
