\chapter{Species}\index{species}\label{species}

\begin{multicols*}{2}
    Species are goupings of characters with dramatic physical differences. In many
    fantasy settings, these are called "races". Examples include humans, dwarves,
    and elves.

    In some settings, species are subdivided into ethnicities. In these cases,
    characters have both a species and an ethnicity. Ethnicities have no mechanical
    effects other than - in some settings - being discriminated against, feared or
    respected.

    During character creation, you choose one species for your character to belong
    to, apply its modifiers, and add its abilities (if any) to your character
    sheet.

    The species listed here-in are a typical selection for a fantasy setting.
    Setting guides will have specific rules for which species you can select
    during character creation.

    \section{Mixed Species}

    Depending on the setting, the GM may allow characters to be a mixed species.

    To create a mixed species, choose one species for each of the character's
    parents. Add all the modifiers together from both parents and divide them
    by two, rounding down to a whole number. Inherit abilities that are shared by
    both parents. If an ability is held by only one parent, inherit the halved
    version from table \ref{tab:inherited-abilities}. Check the glossary for
    any unfamiliar abilities.

    The species a character of mixed species inherits from can be identified
    visually. A character may be discriminated against, feared, or respected based
    on its inherited species depending on the setting.

    A character can also have one or more ancestors of mixed species, but this is
    best represented using the rules above as the mathematics are not granular
    enough to support more specific combinations.
    {
        \unclassedrowcolors
        \begin{tabularx}{0.5\textwidth}{l X}
            \textbf{Ability} & \textbf{Halved Ability} \\
            2 feat points & 1 feat point \\
            Dark vision & Night vision \\
            Small & Slightly Small \\
        \end{tabularx}
        \captionof{table}{Inherited Abilities}
        \label{tab:inherited-abilities}
    }

    \import{./}{human}

    \import{./}{dwarf}

    \import{./}{elf}

    \import{./}{halfling}

    \import{./}{orc}
\end{multicols*}

