\chapter{Character Creation}\label{character-creation}

\begin{multicols}{2}
    To create a character, follow these steps:
    \begin{enumerate}
        \item Distribute your attributes (pg. \pageref{attributes})
        \item Choose your species (pg. \pageref{species})
        \item Choose your culture (pg. \pageref{culture})
        \item Choose your background (pg. \pageref{background})
        \item Distribute your skill points (pg. \pageref{skills})
        \item Choose your feats (pg. \pageref{feats})
        \item Buy your starting equipment (pg. \pageref{equipment})
        \item Choose your starting spells, if applicable (pg. \pageref{magic})
    \end{enumerate}

    \subsection{Advancement}
    % Rewards for a level 1 monster is 60 XP split among the party.
    % Rewards for a level 2 monster is 120 XP split among the party. etc.
    Characters advance by gaining levels, and they gain levels by gaining enough
    experience points (XP). XP is awarded by the GM for overcoming challenges.
    \begin{table*}[ht]
        \unclassedrowcolors
        \begin{tabularx}{\textwidth}{l X l X}
            \textbf{Level} & \textbf{Threshold} & \textbf{Level} & \textbf{Threshold} \\
            1 & 100 & 12 & 2400 \\
            2 & 250 & 13 & 2600 \\
            3 & 450 & 14 & 2800 \\
            4 & 650 & 15 & 3000 \\
            5 & 900 & 16 & 3200 \\
            6 & 1200 & 17 & 3400 \\
            7 & 1400 & 18 & 3600 \\
            8 & 1600 & 19 & 3800 \\
            9 & 1800 & 20 & 4000 \\
            10 & 2000 & 21 & 4200 \\
            11 & 2200 & 22+ & Level $\times$ 200 \\
        \end{tabularx}
        \caption{Levels}
        \label{tab:levels}
    \end{table*}

    Once a character has gained enough XP to meet its current threshold (see table
    \ref{tab:levels}), it becomes able to level up. A character can only level
    up when sleeping. Upon levelling up, a character makes the following changes:
    \begin{itemize}
        \item The character rolls its health die, and then adds the result +
            END (minimum 1) to their maximum health (their health also increases by this
            amount).
        \item The character rolls its mana die, and then adds the result + SOR
            (minimum 0) to their maximum mana (their mana also increases by
            this amount).
        \item The character increases any attribute by 1.
        \item The character increases the potential of X different skills by 1,
            where X is their INT (at least 1).
        \item The character gains 1 feat point.
        \item XP is reduced by the threshold of their current level (see table
            \ref{tab:levels}), and their level is increase by 1.
    \end{itemize}
\end{multicols}
