\chapter{Magic}\index{magic}\label{magic}

\begin{multicols*}{2}
    Magic is a supernatural force accessible in many settings that
    allows characters and creatures to effect the world around them.
    It is the domain of wizards and witches; of mages and sorcerers. If
    you can picture in your mind's eye an old man in a pointy hat
    whispering an incantation and hurling a fireball at his enemies,
    you already have a pretty good idea of what magic is.

    The Unclassed System provides a magic subsystem that should map
    fairly easily onto most fictional settings. It uses a familiar
    resource - mana points (MP)\index{mana points} - governed by an
    attribute\index{attributes} - SOR\index{sorcery} (pg. \label{sorcery}) -
    that expresses your ability to draw on the magical energy of the
    setting. By expending mana points, you can cast spells with
    powerful supernatural effects.

    The magic subsystem is not to be confused with the subsystems for
    religion and psionics. Although these three systems all govern the
    usage of powerful supernatural effects, they have distinct flavors
    and - in the spirit of the Unclassed system - distinct mechanics
    to represent those flavors. While magic draws on the power of
    the world and uses a simple mana-and-spell system common to
    settings with magic, psionics focuses on the ability of the mind
    (and is therefore more ability-based) while religion is predicated
    on your relationship with a specific supernatural force or being
    from which you derive your power. You can envision magic as being
    more knowledge-and-study-based for characters that are more
    INT-focused\index{intelligence} or based on innate magical talent
    for more SOR-focused\index{sorcery} characters.

    Of course, if your GM is running a campaign in a setting with a
    more specific magic system (such as vancian spellcasting), these
    rules will likely prove insufficient. In that case, you should
    ignore this section and instead use a custom magic system, or a
    magic system from a world guide for your setting. Your GM will
    advise you if this is necessary.

    \section{Magic Rules}
    All characters - provided they have at least 1 MP and affordable
    spells to cast - can use magic. They can uses their MP to cast
    spells both in and out of combat, but the particulars of how
    they cast each spell depends on the statistics and rules of the
    spell itself.
    
    \subsection{Mana Points}\label{mana points}\label{maximum mana points}\index{mana points}\index{maximum mana points}
    Mana points (MP) are a well of innate magical energy each character
    has which can be expended to cast spells. A character can have
    a number of MP up to their maximum MP, and can drain their MP as
    low as zero but no further.

    Whenever a character casts a spell, that spell will list an MP
    cost. A character can only cast a spell if they have at least as
    much MP as the cost, in which case they lose that amount of MP
    the instant they begin casting the spell.

    \subsubsection{Recovering Mana}
    Whenever a character rests, they recover all their MP up to their
    maximum MP. You may also encounter various items and abilities
    that can restore MP without having to rest.

    \subsection{Casting Spells}
    Although this section may appear quite large and intimidating,
    casting a spell is ultimately very simple. To cast a spell, after
    expending the required MP, the character makes the skill check for
    the school listed in the spell's description. The spell description
    then explains the effect of the spell, how your skill check
    affects it, and any further rolls you need to make.

    That is all for the general case, but the rest of this section
    covers the specifics and edge cases. Many spell descriptions
    re-iterate these rules in simple terms, and you will only need
    to refer here for the specifics of their terminology.

    The character that casts a spell is referred to as the "caster".

    \subsubsection{Prerequisites}
    Aside from the basic MP cost, there are several other requirements
    for spellcasting. These depend on the spell, and are generally
    listed at the top of the spell description with its other
    statistics.

    \unclassedprop{Gestures}
    {
        Some spells require you to make physical gestures.
        Vocalizations are also included in this category. Any appendage
        (or voice) required by a spell must be otherwise free for the
        casting time of the spell. Taking any other action with an
        appendage (or voice) required by a spell will immediately
        stop casting as if it had been successfully countered.
        Common entries under gestures include "vocal", "one hand",
        and/or "two hands".
    }
    \unclassedprop{Materials}
    {
        Some spells require materials. These are items that the caster
        must hold or manipulate in order to cast the spell. You require
        a free hand to handle materials, but you can also use any hand
        that is being used for gestures. Materials are not consumed
        when the spell is cast.
    }
    \unclassedprop{Materials (Consumed)}
    {
        Occasionally, some spells consume materials when they are cast.
        These materials are handled like normal, but when the spell is
        successfully cast, the material instantly disappears. If the
        spell is countered before casting, you do not waste the
        material, but you also do not get the material back if it is
        countered after casting.
    }
    \unclassedprop{Casting Time}
    {
        Spells take varying amounts of time to cast, and their effects
        only take place once the spell is cast. The most common casting
        time is 1 action, in which case there isn't much to track as
        you simply cast the action and complete the spell during your
        turn, provided you have an action to spend. Longer casting
        times, however, will spill over across rounds. In that case,
        all involved appendages will be occupied until the same moment
        in the turn the specified number of rounds later, at which
        point the spell will finish casting.
    }

    \subsubsection{Skill Check/Spell Power}\label{spell power}\index{skill check}
    The caster makes a skill check of the school corresponding to the
    spell they are casting. The spell description will often call for
    this skill check as a reminder (e.g. "the caster makes an illusion
    skill check") near the beginning of the spell's description. This
    skill check should be noted and retained for the combined casting
    time and duration of the spell, as it becomes the spell power,
    which is used for later calculations.

    A spell's description may state that those characters affected
    by a spell can make a skill check to avoid and/or reduce its
    affects. If so, the target of this skill check will usually be
    the spell power (a.k.a. the caster's skill check).

    \subsubsection{Spell Properties}
    While the description often contains most of what you need to know
    about a spell's effects, there are several properties that are so
    common among spells that they will be listed as properties near
    the beginning of the description for convenient reference.

    \unclassedprop{Duration}
    {Spell effects are applied for the spell's duration. If this is
    instantaneous, they have some instant effect (such as damage),
    and then you can forget about the spell. For longer durations,
    usually some buff is applied until the same time during a turn
    that the spell was cast a certain number of rounds away. If you
    see a spell effect that would usually be instantaneous but it
    lasts for 2 or more rounds, then it is applied at the same time
    every round for that number of rounds (that number includes the
    first instance, which is why 1 round isn't any different from
    instantaneous).} \\
    \unclassedprop{School}
    {The school is the skill that a caster uses in their skill check
    when they cast the spell. For most spells, this is the only
    skill check the caster needs to make, but the full description
    will specify any exceptions.} \\
    \unclassedprop{Spell Power}
    {Usually the spell power is just the caster's skill check, but
    in some cases spells will specify a modifier (typed as a 
    "spell power" modifier) to modify the skill check and therefore
    the spell power.} \\
    \unclassedprop{Damage}
    {For relatively simple damage-dealing spells, the amount of
    damage and the damage type may be listed as a property at the
    start of the spell. All spells have an additional "magical"
    damage type with the "and" relationship, but it usually isn't
    stated in order to save space.} \\
    \unclassedprop{Attack}
    {If a spell is a magic attack, this property will be present
    and will simply read "yes". If it is not present, the spell
    likely has bespoke rules for how it takes effect on targets.} \\
    \unclassedprop{Targets}
    {This property will list the valid targets for the spell. For
    magic attacks, this will likely be melee, ranged, or area. For
    other kinds of spells this will likely be self, touch, ranged,
    or area. Note that for "touch" the valid targets are the same
    as "melee", with the exception that you don't have to make a
    successful magic attack (instead there are likely to be bespoke
    rules about skill contests).} \\
    \unclassedprop{Range}
    {Ranged spells have a range specified in feet. For magic
    attacks, the rules are similar to those for ranged weapons,
    with range not being a hard limit, but rather an incremental
    penalty on the skill check for each multiple of the range the
    target exceeds.} \\
    \unclassedprop{Shape}
    {Area spells will specify the size and shape of the area of
    effect (e.g. "30 ft. sphere" or "50 ft. cube"). For regular
    shapes, the measurement is the diameter or length. More complex
    shapes will have further details.} \\
    \unclassedprop{Spell Effect(s)}
    {The "spell effects" property can generally be ignored during
    normal gameplay. Spells in the Unclassed system are all backed
    by a sophisticated crafting system. This property lets you know
    how the spell was created using the "Designing Spells" rules.
    It is mostly just for interest if the GM or advanced players
    want to reverse engineer spells out-of-game to see how they
    were constructed.}

    \subsubsection{Magic Attacks}
    If the spell's description calls for a magic attack, the caster
    still makes the usual skill check, but as well as setting the spell
    power, the skill check also acts as an attack contested against
    the target's defense check in the usual way (pg.
    \ref{combat:attacking}). % TODO: State that the rules on area attacks are
    % also iterated there (once they are).

    \subsection{Countering Spells}
    Some items, spells, or abilities allow characters to counter a
    spell cast by another character. Countering a spell while it is
    being cast stops the caster's casting short and prevents the spell
    from ever taking effect. Countering a spell after it has been cast,
    during the duration of its effects, will end the spell and any
    effects it is conferring. Either way, the MP cost is not refunded
    to the caster unless allowed by some other ability they have.

    A counter will either specify a spell power - in which case when
    the counter is used any spell that has that spell power or less is
    successfully countered - or call for a skill check. In the latter
    case, if the skill check is greater than or equal to the spell
    power, the spell is successfully countered.

    Unless otherwise stated, unsuccessful spells have no effect on
    the targeted spell.

    \subsection{Perceiving Spells}
    Spells carry immense power, and are thus almost instantly noticed
    by anyone nearby. The casting process often looks and sounds
    extremely dramatic, and consequently the caster's enemies will
    almost certainly want to dispatch the caster quickly.
    
    It is up to the GM whether combat opponents in a setting are wont
    to "gank the wizard" once revealed, but even then, you can imagine
    that there are some spells you might want to cast discretely even
    out of combat.

    It is therefore useful to know the limits of perceiving spells. How
    spells are perceived depends on the details of the spell. Gestures
    make a spell easily perceivable during casting - physical gestures
    by sight, and vocal gestures by sound. But even if you can use
    those gestures, the effects of a spell are still likely to be
    visible. If this is not explicitly stated in the spell, it will
    take some common sense on the part of the GM to make a ruling.
    
    \import{./}{spell-list}
    
    \import{./}{designing-spells}
	
    \import{./}{spell-effect-list}

\end{multicols*}
