The \textbf{cover rating} of a piece of armor represents the difficulty of
finding a weak point and bypassing it. A higher cover rating means a greater \textit{chance}
of reducing damage (see pg. \ref{combat} for how attacking works). The cover
rating of a \textbf{shield} acts as a modifier to your block check (see pg.
\ref{skill:block} for blocking).

The \textbf{Armor Rating} of a piece of armor represents the difficulty of
penetrating it. A higher armor rating means a greated \textit{damage reduction}
when a piece of armor successfully covers you. Shields do not have an armor
rating.

The only penalty from armor is its weight. If you are not strong enough to
carry your armor and all your equipment, you will incur potentially
crippling encumbrance penalties. See the rules on encumbrance (pg.
\ref{encumbrance}).

Table \ref{tab:armor} shows the cost and weight of a full suit of armor. A full
suit of armor is made of a helmet, a cuirass, greaves, a left boot, a right
boot, a left gauntlet, a right gauntlet, a left pauldron and a right pauldron.
All of these pieces have the same cover rating and armor rating as the full
suit. Their cost and weight is a fixed proportion of that of a full suit, as
shown in \ref{tab:armor-pieces}. Do not round.

\begin{table}[h!]
    \begin{tabular}{l l l l l}
        \multicolumn{5}{l}{Suits of Armor} \\
        \bfseries{Armor} & \bfseries{Cost} & \bfseries{Cover Rating} & \bfseries{Armor Rating} & \bfseries{Weight} \\
        Leather & & & & \\
        Plate & 12 & & 16 & \\
        \multicolumn{5}{l}{Shields} \\
        \bfseries{Armor} & \bfseries{Cost} & \bfseries{Cover Rating} & \bfseries{Armor Rating} & \bfseries{Weight} \\
        Light steel shield & & & - & \\
    \end{tabular}
    \caption{Armor}
    \label{tab:armor}
\end{table}

\begin{table}[h!]
    \begin{tabular}{l l l}
        \bfseries{Armor piece} & \bfseries{Cost} & \bfseries{Weight} \\
        Helmet & 0.1$\times$ & 0.1$\times$ \\
        Cuirass & 0.4$\times$ & 0.4$\times$ \\
        Greaves & 0.3$\times$ & 0.3$\times$ \\
        Boot & 0.05$\times$ & 0.05$\times$ \\
        Pauldron & 0.05$\times$ & 0.05$\times$ \\
        Gauntlet & 0.05$\times$ & 0.05$\times$ \\
    \end{tabular}
    \caption{Armor Pieces}
    \label{tab:armor-pieces}
\end{table}
